---
title: "Assignment 4 - Q2"
author: "Mirrien Liang"
documentclass: ctexart
date: "Feb 8, 2022"
output:
  pdf_document:
    latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(global.device = TRUE)


```

```{r results='hide',message=FALSE,warning=FALSE}

library(RSQLite)
library(rworldmap)
library(rworldxtra)
library(MASS)
library(ggmap)

dbcon <- dbConnect(SQLite(),dbname='lab03.sqlite')

```

## Q2a:

```{r}

worldmap = getMap(resolution = "high")
NrthAm = worldmap[which(worldmap$REGION =="North America"),]

plot(NrthAm, xlim=c(-123.35,-122.65),
     ylim=c(49,49.35), main = "Pokemon in Vancouver")


loc1 = "SELECT latitude,longitude FROM Vanpoke"
location = dbGetQuery(dbcon, loc1)

points(location$longitude,location$latitude,cex=0.5,col="gray35")
```

## Q2b:

```{r}
f1 <- kde2d(location$longitude, location$latitude,n=200)

contour(f1$x,f1$y,f1$z,
        drawlabels=FALSE,
        nlevels=10,
        col=c("black","darkblue","darkgreen","darkred","red"),
        add=TRUE,
        lwd = 2)
```

### Reference:

https://stackoverflow.com/questions/37899785/contour-plot-density-of-sites-with-latitude-and-longitude-locations

https://stats.stackexchange.com/questions/31726/scatterplot-with-contour-heat-overlay

https://www.r-bloggers.com/2014/09/5-ways-to-do-2d-histograms-in-r/

```{r include=FALSE}
knitr::opts_knit$set(global.device = FALSE)
```


## Q2c:

In three sentences or fewer, answer these questions: Where are the peaks of this two dimensional density plot? Why are the peaks in those locations?

According to the density plot in the previous question, based on the color of contour lines and a general presentation of points, the peaks appear at (1) Vancouver City Hall, (2) Burnaby City Hall, (3) somewhere around the City of Lougheed Shopping Centre, and (4) somewhere around Lonsdale Quay Market in North Vancouver.

## Q2d:

If we wish to depict Vancouver in greater details, *"ggmap"* allows us to import a satellite stamen map of Vancouver areas and to plot points and layers on it.
Along with *"ggplot"* as a powerful tool, it helps us generate vivid density plots and contour lines/areas over the map.

In the following chunk of codes, I use *"qmplot"* to plot Pokemon location scatter points on the map imported. By using *"geom_density_2d"* and *"stat_density_2d"*, I am able to overlay the contour lines and filled areas over the scatter plot. The last function is to adjust the color palette of the areas between contour lines.


The codes and the resulted graph are presented as follows:

```{r results='hide',message=FALSE,fig.keep='all'}
qmplot(x=longitude, y=latitude,
       data = location,
       maptype = "toner-lite",
       color = I("black"),
       alpha = I(.2)) +
  geom_density_2d(color = "black") +
  stat_density_2d(aes(fill = ..level..),
                  geom = "polygon",
                  alpha = .4,
                  color = NA,
                  contour = TRUE) +
  scale_fill_gradient2("Density",
                       low = "white",
                       mid = "yellow",
                       high = "red",
                       midpoint = 10)

```

### Reference:

https://www.r-graph-gallery.com/2d-density-plot-with-ggplot2.html

https://cran.r-project.org/web/packages/ggmap/readme/README.html

https://rpubs.com/nickbearman/r-google-map-making

https://biostats.w.uib.no/creating-a-2d-density-plot/